---
title: Abigail I. Gutierrez 1.4 
output: html_notebook
---

Part 1: Mapping 
[BIEN](http://biendata.org) 

Mapping and Downloading Data. 

```{r}
Castanea_dentata <- BIEN_occurrence_species(species = "Castanea dentata")  
Castanea_dentata 

str(Castanea_dentata)
head(Castanea_dentata)
tail (Castanea_dentata) 

```

Now want to get range in map of where the Chestnut Tree resides 
```{r}
Castanea_dentata_rangemap <- BIEN_ranges_load_species("Castanea dentata")

Castanea_dentata_rangemap 

# Map the range of chestnut tree
map('world', fill=T, col="grey", bg = "light blue", xlim=c(-180,20),ylim=c(-60,80))
plot(Castanea_dentata_rangemap, col="green", add=T) 

```
Now mapping range of occurance 
```{r}
# Map the range of chestnut tree + occurence records
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
plot(Castanea_dentata_rangemap, col="green", add=T)
points(cbind(Castanea_dentata$longitude, Castanea_dentata$latitude), col="blue", pch=20, cex=1) 
```

Getting more expansive data using more information and expanding the map to the entire world 

```{r}

Castanea_dentata_full <- BIEN_occurrence_species(species = "Castanea dentata",cultivated = T,only.new.world = T,all.taxonomy = T,native.status = T,observation.type = T,political.boundaries = T)

Castanea_dentata_full 
str(Castanea_dentata_full) 

```
 World Map 
 The blue defines the original data set and the red is from all of the occurences. 
 
```{r}
map('world', fill=T , col="grey", bg="light blue")

#Plot the points from the full query in red

points(cbind(Castanea_dentata_full$longitude, Castanea_dentata_full$latitude), col="red", pch=20, cex=1)

# Plot the points from the default query in blue
points(cbind(Castanea_dentata$longitude, Castanea_dentata$latitude), col="blue", pch=20, cex=1) 
```
 Now attempting for a second species. This time I want to do the Common pricklyash. 
```{r}
Zanthoxylum_americanum <- BIEN_occurrence_species(species = "Zanthoxylum americanum")  
Zanthoxylum_americanum

str(Zanthoxylum_americanum)
head(Zanthoxylum_americanum)
tail (Zanthoxylum_americanum)  
```
 Now want to get range in map of where the Common pricklyash resides  
```{r}
Zanthoxylum_americanum_rangemap <- BIEN_ranges_load_species("Zanthoxylum americanum")

Zanthoxylum_americanum_rangemap 

# Map the range of aok tree
map('world', fill=T, col="grey", bg = "light blue", xlim=c(-180,20),ylim=c(-60,80))
plot(Zanthoxylum_americanum_rangemap, col="green", add=T)  
```
 Now we want to get range of occurance 
```{r}
# Map the range of chestnut tree + occurence records
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
plot(Zanthoxylum_americanum_rangemap, col="green", add=T)
points(cbind(Zanthoxylum_americanum$longitude, Zanthoxylum_americanum$latitude), col="blue", pch=20, cex=1)  
```
 Getting more expansive data using more information and expanding the map to the entire world 
```{r}
Zanthoxylum_americanum_full <- BIEN_occurrence_species(species = "Zanthoxylum americanum",cultivated = T,only.new.world = T,all.taxonomy = T,native.status = T,observation.type = T,political.boundaries = T)

Zanthoxylum_americanum_full 
str(Zanthoxylum_americanum_full)  
```

World Map 
The blue defines the original data set and the red is from all of the occurences.  
 
```{r}
map('world', fill=T , col="grey", bg="light blue")

#Plot the points from the full query in red

points(cbind(Zanthoxylum_americanum_full$longitude, Zanthoxylum_americanum_full$latitude), col="red", pch=20, cex=1)

# Plot the points from the default query in blue
points(cbind(Zanthoxylum_americanum$longitude, Zanthoxylum_americanum$latitude), col="blue", pch=20, cex=1) 

```
 
 Part 2: Trait Data 
```{r}
BIEN_trait_list() 
```
Creating a data frame of American Chestnut Tree Traits 

```{r}
Castanea_dentata_traits <- BIEN_trait_genus(genus = "Castanea")
str(Castanea_dentata_traits)
Castanea_dentata_traits
```
```{r}
Castanea_dentata_WPH <- BIEN_trait_traitbygenus(trait = "whole plant height", genus = "Castanea")
head(Castanea_dentata_WPH)
tail(Castanea_dentata_WPH)
str(Castanea_dentata_WPH)

# List variable names from Breadfruit_DBH data frame
names(Castanea_dentata_WPH)

Castanea_dentata_WPH_small = Castanea_dentata_WPH[c(1:5)]
Castanea_dentata_WPH_small
Castanea_dentata_WPH_small = as.data.frame(Castanea_dentata_WPH_small)
class(Castanea_dentata_WPH_small)
head(Castanea_dentata_WPH_small)
class(Castanea_dentata_WPH_small$trait_value)
WPH = as.numeric(Castanea_dentata_WPH_small$trait_value)

class(WPH) 

```
Boxplot 
```{r}
#now we will be doing a boxplot

library(lattice)
head(Castanea_dentata_WPH_small)
boxplot(WPH~scrubbed_species_binomial, data=Castanea_dentata_WPH_small) 
```

 Creating another frame for a different trait. 
```{r}
Castanea_dentata_DBH <- BIEN_trait_traitbygenus(trait = "diameter at breast height (1.3 m)", genus = "Castanea")
str(Castanea_dentata_DBH)
Castanea_dentata_DBH


head(Castanea_dentata_DBH)
tail(Castanea_dentata_DBH)
str(Castanea_dentata_DBH)

# List variable names from Breadfruit_DBH data frame
names(Castanea_dentata_DBH)

Castanea_dentata_DBH_small = Castanea_dentata_DBH[c(1:5)]
Castanea_dentata_DBH_small
Castanea_dentata_DBH_small = as.data.frame(Castanea_dentata_DBH_small)
class(Castanea_dentata_DBH_small)
head(Castanea_dentata_DBH_small)
class(Castanea_dentata_DBH_small$trait_value)
DBH = as.numeric(Castanea_dentata_DBH_small$trait_value)

class(DBH) 

head(Castanea_dentata_DBH_small)
boxplot(DBH~scrubbed_species_binomial, data=Castanea_dentata_DBH_small)  

```
 
 Summary Statistics: 
 
 For this assignment I decided to look at whole plant height and diameter of breast height at 1.3m for Castanea_dentata and I looked at the trait from the genus Castanea. The mean of the diameter of breast height at 1.3m I got a mean of 5.777cm . However for the mean of whole plant height I got a value of 7.608m. There is a 100 fold difference since we are talking about values that differ in m and cm.   
 
```{r}
mean(DBH, data=Castanea_dentata_DBH_small)

mean(WPH, data=Castanea_dentata_WPH_small)

```
 
 Now I want to do a scatter plot and a histogram using the same traits I have used so far 
```{r}
#whp stands for whole plant height 
boxplot(WHP)
hist(WPH) 

#DBH is diameter at breast height 
boxplot(DBH)
hist(DBH) 

```
 Now I want to attempt to run a t-test on the data I collected on the whole plant height. I tried to run the test on the trait of whole plant height but I was unable to be 
```{r}

ttest_result1 = t.test(x=Castanea_dentata_WPH_small$scrubbed_species_binomial, y=Castanea_dentata_WPH_small$trait_value, paired=TRUE, alternative="two.sided")
ttest_result1 
```
 
Great job, Abby! All of your code worked perfectly until the very end on line 195, but that was only because you typed "WHP" instead of "WPH" following the boxplot() function.

Sorry for the t-test frustration. Did you see Hilary's announcement about it on Canvas? https://courseworks2.columbia.edu/courses/76272/discussion_topics/310707
For full credit, go follow those instructions and then push your code back to GitHub once the t-test part works.